---
title: "Inferential Feature Analysis: Latino Trump Support (CMPS 2020)"
subtitle: "Analysis Pipeline — Random Forest + SHAP + Bootstrap"
author: "J. Grijalva"
date: today
format:
  pdf:
    toc: true
    toc-depth: 3
    number-sections: true
    fontsize: 11pt
    geometry: margin=1in
    fig-pos: "H"
    keep-tex: false
execute:
  echo: false
  message: false
  warning: false
  cache: false
editor: source
params:
  fast_dev: false    # PRODUCTION RENDER: 500 bootstraps, 500 trees. Flip to true for 10-boot / 100-tree dry run.
  n_trees: 500
  n_boot: 500
  prop_test: 0.25
  raw_data_file: "39096-0001-Data.rda"
---

```{r}
#| label: setup

required_pkgs <- c(
  "tidyverse", "ranger", "treeshap", "shapviz", "missRanger",
  "fastDummies", "caret", "pROC", "glmnet", "xgboost",
  "ggcorrplot", "knitr", "kableExtra", "sessioninfo"
)

missing_pkgs <- required_pkgs[!vapply(required_pkgs, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing_pkgs) > 0) {
  stop(
    "Missing required packages: ", paste(missing_pkgs, collapse = ", "),
    "\nInstall them before rendering.",
    call. = FALSE
  )
}

invisible(lapply(required_pkgs, library, character.only = TRUE))

options(width = 120)
knitr::opts_chunk$set(
  echo      = FALSE,
  message   = FALSE,
  warning   = FALSE,
  cache     = FALSE,
  dpi       = 300,
  fig.align = "center"
)
```

```{r}
#| label: constants

FAST_DEV   <- isTRUE(params$fast_dev)
N_TREES    <- as.integer(params$n_trees)
N_BOOT     <- if (FAST_DEV) 10L else as.integer(params$n_boot)
PROP_TEST  <- as.numeric(params$prop_test)
WEIGHT_VAR <- "WEIGHT"

# Keep FALSE if you want continuity with your previously run 2020 pipeline.
# Flip to TRUE if you want full parity with the 2016 weighting logic.
USE_SURVEY_WEIGHTS <- TRUE

# OHE choice retained from the 2020 pipeline to keep the encoded feature space
# tractable for benchmark models.
REMOVE_FIRST_DUMMY <- TRUE

# Decision rule: feature j is a finding if ci_lo(j) > 0 AND sigma_j <= SIGMA_THRESHOLD
# Pre-specified before examining results. Reported alongside all findings.
SIGMA_THRESHOLD <- 3.0

# Reproducibility seeds — fixed, never change
SEED_SPLIT      <- 2026L
SEED_IMPUTE     <- 2027L
SEED_RF_T1      <- 2028L
SEED_RF_T2      <- 2029L
SEED_RF_T3      <- 2030L
SEED_RF_MEN     <- 2031L
SEED_RF_WOMEN   <- 2032L
SEED_BOOT       <- 2033L
# Note: SEED_BOOT + 1000L and + 2000L are reserved as T1 / T2 bootstrap offsets
# (matches 2016 / 2024 convention). The gender-subgroup QMD seeds its own bootstraps
# separately and does not collide with these.

# Locate raw data
PROJECT_ROOT <- normalizePath(getwd(), winslash = "/", mustWork = FALSE)
RAW_DATA_CANDIDATES <- unique(c(
  params$raw_data_file,
  file.path(PROJECT_ROOT, "data", "raw", "39096-0001-Data.rda"),
  file.path(PROJECT_ROOT, "data", "39096-0001-Data.rda"),
  file.path(PROJECT_ROOT, "39096-0001-Data.rda"),
  file.path(PROJECT_ROOT, "data", "raw", "ICPSR_39096", "DS0001", "39096-0001-Data.rda")
))
RAW_DATA_CANDIDATES <- RAW_DATA_CANDIDATES[nzchar(RAW_DATA_CANDIDATES)]
RAW_DATA_FILE <- RAW_DATA_CANDIDATES[file.exists(RAW_DATA_CANDIDATES)][1]
if (is.na(RAW_DATA_FILE)) {
  stop(
    "Raw data file not found. Searched:\n",
    paste0("- ", RAW_DATA_CANDIDATES, collapse = "\n"),
    call. = FALSE
  )
}
```

```{r}
#| label: helpers

# --- Weight rescaling ---
rescale_w <- function(w) {
  w <- as.numeric(w)
  w[!is.finite(w)] <- NA_real_
  if (all(is.na(w))) stop("Weight vector is entirely missing.", call. = FALSE)
  w[is.na(w)] <- median(w, na.rm = TRUE)
  w / mean(w)
}

# --- Expand column names matching stems ---
expand_drop <- function(cols, stems) {
  unique(unlist(lapply(stems, function(s) {
    grep(paste0("^", s, "(_|$)"), cols, value = TRUE)
  })))
}

# --- 2020 feature labels ---
feature_labels <- c(
  "Q130_(2) The coronavirus is unfortunate, but there is nothi" = "COVID: Trump not to blame",
  "voter_fraud_belief_(5) No, I don"                            = "Voter fraud: definitely no fraud",
  "Q54_(3) Trump had no connection to the rioters, he should n" = "Jan 6: Trump not responsible",
  "Q684_(2) Some people voted in the election even though they" = "Belief: ineligible voters cast ballots",
  "Q53_(2) A coordinated act of insurrection against the Unite" = "Jan 6: not an insurrection",
  "Q55_(2) Protecting democracy"                                = "Jan 6: protecting democracy",
  "Q56_(3) They were not White supremacists, racism had nothin" = "Jan 6 rioters: not White supremacists",
  "Q575_Q581R2"                                                 = "Feeling thermometer: Rep. AOC [EXCLUDED List D]",
  "Q358_(5) Strongly oppose"                                    = "Oppose BLM-related policy",
  "Q359R1_(5) Strongly disagree"                                = "Racial justice statement 1: disagree",
  "Q359R2_(5) Strongly disagree"                                = "Racial justice statement 2: disagree",
  "Q485_(2) Both the treatment of Blacks by the police and tre" = "Reject intersectional bias framing",
  "Q543_(2) Oppose government increasing spending by $25 billi" = "Oppose border wall spending",
  "Q1R19_(1) Border security to control immigration"            = "Priority: border security",
  "Q1R17_(1) Stopping the left-wing Democratic agenda"          = "Priority: stop left-wing agenda",
  "Q131R2_(5) Strongly oppose"                                  = "Oppose national mask mandate",
  "Q131R6_(5) Strongly oppose"                                  = "Oppose police reform for African Americans",
  "Q131R10_(5) Strongly oppose"                                 = "Oppose DC statehood",
  "Q706_Q708R1_(5) Strongly oppose"                             = "Oppose defunding police departments",
  "Q504R6"                                                      = "Stress: white supremacy",
  "Q504R3"                                                      = "Stress: climate change",
  "Q183_(4) Very favorable"                                     = "ICE/Border Patrol: very favorable",
  "Q685_(1) Domestic threats"                                   = "Greater threat: domestic",
  "Q308R1"                                                      = "Feeling therm: undocumented immigrants",
  "Q308R2"                                                      = "Feeling therm: gun owners",
  "Q308R3"                                                      = "Feeling therm: veterans",
  "Q308R4"                                                      = "Feeling therm: welfare recipients",
  "Q308R5"                                                      = "Feeling therm: climate change activists",
  "Q308R6"                                                      = "Feeling therm: MeToo movement",
  "Q308R7"                                                      = "Feeling therm: Black Lives Matter",
  "Q563_Q572R5"                                                 = "Group representation evaluation",
  "Q225R14"                                                     = "White nationalists threaten society"
)

label_feature <- function(x) {
  out <- x

  # Exact matches first
  exact_hits <- match(out, names(feature_labels), nomatch = 0L)
  out[exact_hits > 0L] <- unname(feature_labels[exact_hits[exact_hits > 0L]])

  # Prefix matches for variants not covered exactly
  out[str_detect(x, "^Q130")]         <- "COVID: Trump not to blame"
  out[str_detect(x, "^voter_fraud_belief")] <- "Voter fraud belief"
  out[str_detect(x, "^jan6_fraud_belief")]  <- "Jan 6 fraud belief"
  out[str_detect(x, "^Q54")]          <- "Jan 6: Trump not responsible"
  out[str_detect(x, "^Q684")]         <- "Belief: ineligible voters cast ballots"
  out[str_detect(x, "^Q53")]          <- "Jan 6: not an insurrection"
  out[str_detect(x, "^Q55")]          <- "Jan 6: protecting democracy"
  out[str_detect(x, "^Q56")]          <- "Jan 6 rioters: not White supremacists"
  out[str_detect(x, "^Q543")]         <- "Border wall spending attitude"
  out[str_detect(x, "^Q358")]         <- "Oppose BLM-related policy"
  out[str_detect(x, "^Q359R1")]       <- "Racial justice statement 1"
  out[str_detect(x, "^Q359R2")]       <- "Racial justice statement 2"
  out[str_detect(x, "^Q485")]         <- "Reject intersectional bias framing"
  out[str_detect(x, "^Q131R2")]       <- "Oppose national mask mandate"
  out[str_detect(x, "^Q131R6")]       <- "Oppose police reform for African Americans"
  out[str_detect(x, "^Q131R8")]       <- "Oppose bold climate policy agenda"
  out[str_detect(x, "^Q131R10")]      <- "Oppose DC statehood"
  out[str_detect(x, "^Q706_Q708R1")]  <- "Oppose defunding police departments"
  out[str_detect(x, "^Q183")]         <- "ICE/Border Patrol favorability"
  out[str_detect(x, "^Q685")]         <- "Perceived threat source"
  out[str_detect(x, "^Q1R17")]        <- "Priority: stop left-wing agenda"
  out[str_detect(x, "^Q1R19")]        <- "Priority: border security"
  out[str_detect(x, "^Q504R4")]       <- "Stress: police brutality"
  out[str_detect(x, "^Q308R1")]       <- "Feeling therm: undocumented immigrants"
  out[str_detect(x, "^Q308R2")]       <- "Feeling therm: gun owners"
  out[str_detect(x, "^Q308R3")]       <- "Feeling therm: veterans"
  out[str_detect(x, "^Q308R4")]       <- "Feeling therm: welfare recipients"
  out[str_detect(x, "^Q308R5")]       <- "Feeling therm: climate activists"
  out[str_detect(x, "^Q308R6")]       <- "Feeling therm: MeToo"
  out[str_detect(x, "^Q308R7")]       <- "Feeling therm: BLM"
  out[str_detect(x, "^Q225R14")]      <- "White nationalists threaten society"
  out
}

# --- LaTeX table helper ---
latex_tbl <- function(df, caption = NULL, font_size = 9, scale_down = FALSE) {
  out <- knitr::kable(df, format = "latex", booktabs = TRUE, linesep = "", caption = caption)
  opts <- if (scale_down) c("scale_down", "hold_position") else c("hold_position")
  kableExtra::kable_styling(out, font_size = font_size, latex_options = opts)
}

# --- MissRanger wrapper ---
impute_missranger <- function(df, seed) {
  set.seed(seed)
  tryCatch(
    missRanger::missRanger(df, num.trees = 100L, verbose = 0, pmm.k = 3L),
    error = function(e) {
      message("missRanger pmm.k=3 failed; retrying with pmm.k=0. Reason: ", conditionMessage(e))
      set.seed(seed)
      missRanger::missRanger(df, num.trees = 100L, verbose = 0, pmm.k = 0L)
    }
  )
}

# --- Numeric cleanup ---
clean_matrix <- function(df) {
  as.data.frame(lapply(df, function(v) {
    v <- as.numeric(v)
    v[is.infinite(v) | is.nan(v)] <- NA_real_
    if (anyNA(v)) {
      m <- median(v, na.rm = TRUE)
      v[is.na(v)] <- if (is.finite(m)) m else 0
    }
    v
  }))
}

# --- OHE + alignment ---
ohe_align <- function(train_df, test_df, remove_first_dummy = TRUE) {
  cat_cols <- names(train_df)[vapply(train_df, function(x) is.factor(x) || is.character(x), logical(1))]
  cat_cols <- intersect(cat_cols, names(test_df))

  if (length(cat_cols) > 0) {
    train_ohe <- fastDummies::dummy_cols(
      train_df,
      select_columns = cat_cols,
      remove_first_dummy = remove_first_dummy,
      remove_selected_columns = TRUE,
      ignore_na = FALSE
    )
    test_ohe <- fastDummies::dummy_cols(
      test_df,
      select_columns = cat_cols,
      remove_first_dummy = remove_first_dummy,
      remove_selected_columns = TRUE,
      ignore_na = FALSE
    )
  } else {
    train_ohe <- train_df
    test_ohe  <- test_df
  }

  missing_cols <- setdiff(names(train_ohe), names(test_ohe))
  extra_cols   <- setdiff(names(test_ohe), names(train_ohe))

  if (length(missing_cols) > 0) {
    for (col in missing_cols) test_ohe[[col]] <- 0
  }
  if (length(extra_cols) > 0) {
    test_ohe <- test_ohe[, !names(test_ohe) %in% extra_cols, drop = FALSE]
  }
  test_ohe <- test_ohe[, names(train_ohe), drop = FALSE]

  list(train = train_ohe, test = test_ohe,
       missing_cols = missing_cols, extra_cols = extra_cols)
}

# --- Prepare train/test matrices for one split ---
prepare_split_matrices <- function(df, feature_pool, train_idx, weights_full,
                                   remove_first_dummy = TRUE, impute_seed = SEED_IMPUTE) {
  x_train_raw <- df[train_idx, feature_pool, drop = FALSE]
  x_test_raw  <- df[-train_idx, feature_pool, drop = FALSE]

  y_train <- df$trump_vote[train_idx]
  y_test  <- df$trump_vote[-train_idx]

  is_male_train   <- df$.is_male[train_idx]
  is_female_train <- df$.is_female[train_idx]
  is_male_test    <- df$.is_male[-train_idx]
  is_female_test  <- df$.is_female[-train_idx]

  w_train <- if (USE_SURVEY_WEIGHTS) rescale_w(weights_full[train_idx]) else rep(1, length(train_idx))
  w_test  <- if (USE_SURVEY_WEIGHTS) rescale_w(weights_full[-train_idx]) else rep(1, nrow(df) - length(train_idx))

  x_train_raw <- x_train_raw |> mutate(across(where(is.character), as.factor))
  x_test_raw  <- x_test_raw  |> mutate(across(where(is.character), as.factor))

  all_na_cols <- names(x_train_raw)[colSums(is.na(x_train_raw)) == nrow(x_train_raw)]
  if (length(all_na_cols) > 0) {
    x_train_raw <- x_train_raw[, !names(x_train_raw) %in% all_na_cols, drop = FALSE]
    x_test_raw  <- x_test_raw[, !names(x_test_raw) %in% all_na_cols, drop = FALSE]
  }
  x_test_raw <- x_test_raw[, names(x_train_raw), drop = FALSE]

  x_train_imp <- impute_missranger(x_train_raw, impute_seed)
  x_test_imp  <- impute_missranger(x_test_raw,  impute_seed + 1L)

  aligned <- ohe_align(x_train_imp, x_test_imp, remove_first_dummy = remove_first_dummy)
  x_train_num <- clean_matrix(aligned$train)
  x_test_num  <- clean_matrix(aligned$test)

  all_na_cols2 <- names(x_train_num)[colSums(is.na(x_train_num)) == nrow(x_train_num)]
  if (length(all_na_cols2) > 0) {
    x_train_num <- x_train_num[, !names(x_train_num) %in% all_na_cols2, drop = FALSE]
    x_test_num  <- x_test_num[, !names(x_test_num) %in% all_na_cols2, drop = FALSE]
  }

  nzv_idx <- caret::nearZeroVar(x_train_num, saveMetrics = FALSE)
  if (length(nzv_idx) > 0) {
    x_train_num <- x_train_num[, -nzv_idx, drop = FALSE]
    x_test_num  <- x_test_num[, -nzv_idx, drop = FALSE]
  }
  x_test_num <- x_test_num[, names(x_train_num), drop = FALSE]

  list(
    x_train_imp = x_train_num,
    x_test_imp = x_test_num,
    y_train = y_train,
    y_test = y_test,
    w_train = w_train,
    w_test = w_test,
    is_male_train = is_male_train,
    is_female_train = is_female_train,
    is_male_test = is_male_test,
    is_female_test = is_female_test,
    n_missing_cols = length(aligned$missing_cols),
    n_extra_cols = length(aligned$extra_cols)
  )
}

# --- Tier construction ---
build_tier_sets <- function(x_train_imp, x_test_imp, exclude_tautological, exclude_partisan) {
  all_cols <- names(x_train_imp)

  t2_drop <- expand_drop(all_cols, exclude_tautological)
  ring1_drop <- expand_drop(all_cols, exclude_partisan)
  t3_drop <- union(t2_drop, ring1_drop)

  t2_cols <- setdiff(all_cols, t2_drop)
  t3_cols <- setdiff(all_cols, t3_drop)

  partisan_guard_stems <- unique(c(exclude_partisan, "Q14"))
  partisan_leakage <- expand_drop(t3_cols, partisan_guard_stems)
  if (length(partisan_leakage) > 0) {
    stop(
      "Ring 1 leakage audit FAILED. Leaked variables: ",
      paste(partisan_leakage, collapse = ", "),
      call. = FALSE
    )
  }

  list(
    train_t1 = x_train_imp,
    test_t1 = x_test_imp,
    train_t2 = x_train_imp[, t2_cols, drop = FALSE],
    test_t2 = x_test_imp[, t2_cols, drop = FALSE],
    train_t3 = x_train_imp[, t3_cols, drop = FALSE],
    test_t3 = x_test_imp[, t3_cols, drop = FALSE],
    t2_drop = t2_drop,
    t3_drop = t3_drop,
    t2_cols = t2_cols,
    t3_cols = t3_cols
  )
}

# --- Model fitting ---
fit_rf <- function(x, y, seed, survey_w = NULL) {
  x <- as.data.frame(lapply(x, as.numeric))
  y_fac <- factor(y, levels = c(0, 1))
  n0 <- sum(y == 0)
  n1 <- sum(y == 1)
  n <- n0 + n1
  class_w <- c("0" = n / (2 * n0), "1" = n / (2 * n1))
  if (is.null(survey_w)) survey_w <- rep(1, length(y))
  survey_w <- rescale_w(survey_w)
  set.seed(seed)
  ranger::ranger(
    x = x,
    y = y_fac,
    num.trees = N_TREES,
    mtry = floor(sqrt(ncol(x))),
    min.node.size = 5L,
    case.weights = survey_w,
    class.weights = class_w,
    probability = TRUE,
    importance = "impurity",
    seed = seed
  )
}

predict_prob <- function(model, newdata) {
  predict(model, data = as.data.frame(lapply(newdata, as.numeric)))$predictions[, "1"]
}

get_auc <- function(model, x_te, y_te) {
  pred <- predict_prob(model, x_te)
  as.numeric(pROC::auc(pROC::roc(y_te, pred, quiet = TRUE)))
}

# --- Weighted mean absolute SHAP ---
wmean_abs_shap <- function(shap_mat, w) {
  shap_mat <- as.matrix(shap_mat)
  apply(abs(shap_mat), 2, function(col) weighted.mean(col, w = w))
}

# --- SHAP computation ---
# Harmonized with 2024 / 2016: unifier is built on the TRAIN frame (model's fit
# domain) and then applied to the test frame for attribution. Using the test
# frame for the unifier can produce unstable attributions if test-only values
# fall outside the tree's learned split geometry.
compute_shap <- function(rf_model, x_tr, x_te) {
  x_tr_num <- as.data.frame(lapply(x_tr, as.numeric))
  x_te_num <- as.data.frame(lapply(x_te, as.numeric))
  if (sum(is.na(x_tr_num)) > 0) x_tr_num[is.na(x_tr_num)] <- 0
  if (sum(is.na(x_te_num)) > 0) x_te_num[is.na(x_te_num)] <- 0
  unified <- treeshap::ranger.unify(rf_model, x_tr_num)
  ts      <- treeshap::treeshap(unified, x_te_num)
  shapviz::shapviz(ts)
}

# --- Bootstrap wrapper ---
run_bootstrap <- function(train_d, test_d, y_tr, y_te, w_tr, w_te, seed_base, label = "T3") {
  p <- ncol(train_d)
  n <- nrow(train_d)
  b_imp  <- matrix(NA_real_, nrow = N_BOOT, ncol = p, dimnames = list(NULL, names(train_d)))
  b_rank <- matrix(NA_real_, nrow = N_BOOT, ncol = p, dimnames = list(NULL, names(train_d)))
  b_auc  <- rep(NA_real_, N_BOOT)
  x_te_num <- as.data.frame(lapply(test_d, as.numeric))
  if (sum(is.na(x_te_num)) > 0) x_te_num[is.na(x_te_num)] <- 0
  failed <- integer(0)

  for (i in seq_len(N_BOOT)) {
    result <- tryCatch({
      set.seed(seed_base + i)
      idx <- sample.int(n, size = n, replace = TRUE)
      xb <- train_d[idx, , drop = FALSE]
      xb_num <- as.data.frame(lapply(xb, as.numeric))
      if (sum(is.na(xb_num)) > 0) xb_num[is.na(xb_num)] <- 0
      rf_b <- fit_rf(xb, y_tr[idx], seed = seed_base + i, survey_w = w_tr[idx])
      auc_b <- get_auc(rf_b, test_d, y_te)
      unified <- treeshap::ranger.unify(rf_b, xb_num)
      ts_b <- treeshap::treeshap(unified, x_te_num)
      imp_b <- wmean_abs_shap(ts_b$shaps, w_te)
      list(auc = auc_b, imp = imp_b)
    }, error = function(e) {
      message("[", label, "] Bootstrap iteration ", i, " failed: ", conditionMessage(e))
      NULL
    })

    if (is.null(result)) {
      failed <- c(failed, i)
      next
    }

    b_auc[i] <- result$auc
    b_imp[i, ] <- result$imp
    b_rank[i, ] <- rank(-result$imp, ties.method = "average")

    if (i %% 25 == 0 || i == N_BOOT) {
      message("[", label, "] Completed iteration ", i, " / ", N_BOOT)
    }
  }

  good <- !is.na(b_auc)
  if (sum(good) < 10) {
    stop("[", label, "] Fewer than 10 successful bootstrap iterations.", call. = FALSE)
  }
  if (length(failed) > 0) {
    message("[", label, "] ", length(failed), " iterations failed and were dropped.")
  }

  list(
    imp = b_imp[good, , drop = FALSE],
    rank = b_rank[good, , drop = FALSE],
    auc = b_auc[good],
    n_success = sum(good),
    n_failed = length(failed)
  )
}

# --- AUC CI text helper ---
auc_ci <- function(boot_auc) {
  paste0("[", round(quantile(boot_auc, 0.025, na.rm = TRUE), 3), ", ",
         round(quantile(boot_auc, 0.975, na.rm = TRUE), 3), "]")
}
```

# Analysis Parameters

```{r}
#| label: tbl-params
#| tbl-cap: "Analysis Configuration."

tibble(
  Parameter = c("Bootstrap iterations (B)", "Trees per forest",
                "Test proportion", "Rank stability threshold (k)",
                "Development mode", "Raw data file"),
  Value = c(N_BOOT, N_TREES, PROP_TEST, SIGMA_THRESHOLD,
            FAST_DEV, basename(RAW_DATA_FILE))
) |>
  latex_tbl(font_size = 9)
```

# Data and Sample Construction

```{r}
#| label: data-load
#| include: false

e <- new.env(parent = emptyenv())
load(RAW_DATA_FILE, envir = e)
nms <- ls(e, all.names = TRUE)
if ("da39096.0001" %in% nms) {
  cmps_raw <- get("da39096.0001", envir = e)
} else {
  dfs <- nms[vapply(nms, function(x) is.data.frame(get(x, envir = e)), logical(1))]
  cmps_raw <- get(dfs[1], envir = e)
}
rm(e)

cmps_latino <- cmps_raw |>
  mutate(.row_id = seq_len(n())) |>
  filter(
    str_detect(as.character(MAIN_ETH_QUOTA), "^\\(2\\)") |
      str_detect(as.character(MAIN_ETH_QUOTA), "[Hh]ispanic|[Ll]atino")
  ) |>
  filter(str_detect(as.character(Q12), "^\\(1\\)|^\\(2\\)")) |>
  mutate(
    trump_vote = case_when(
      str_detect(as.character(Q14), "^\\(1\\)") ~ 1L,
      is.na(Q14)                                   ~ NA_integer_,
      TRUE                                         ~ 0L
    )
  ) |>
  filter(!is.na(trump_vote))

# --- CRITICAL: Fix gender labels from raw data BEFORE any preprocessing ---
# Gender membership is a fixed sample property, never derived from imputed features.
gender_raw <- as.character(cmps_latino$S3B)
cmps_latino$.is_male   <- as.integer(str_detect(gender_raw, "^\\(1\\)"))
cmps_latino$.is_female <- as.integer(str_detect(gender_raw, "^\\(2\\)"))

weights_full <- cmps_latino[[WEIGHT_VAR]]
if (is.null(weights_full)) weights_full <- rep(1, nrow(cmps_latino))
stopifnot(nrow(cmps_latino) > 0, sum(cmps_latino$trump_vote) > 0)
```

```{r}
#| label: tbl-sample
#| tbl-cap: "Latino Subsample Summary."

tibble(
  Item = c("Latino N", "Trump voters", "Trump %", "Weight range"),
  Value = c(
    nrow(cmps_latino),
    sum(cmps_latino$trump_vote),
    paste0(round(mean(cmps_latino$trump_vote) * 100, 1), "%"),
    paste0("[", round(min(weights_full, na.rm = TRUE), 2), ", ", round(max(weights_full, na.rm = TRUE), 2), "]")
  )
) |>
  latex_tbl(font_size = 9)
```

# Preprocessing

```{r}
#| label: preprocess
#| include: false

sweep_na <- function(df) {
  dk_regex <- paste(
    c("don.?t know", "\\bdk\\b", "refused", "not applicable", "\\bn/a\\b", "not sure", "^\\s*$"),
    collapse = "|"
  )
  sentinels <- c(88, 99, 999, 9999)

  df |>
    mutate(
      across(where(is.factor), ~ {
        lvls <- levels(.)
        lvls <- iconv(lvls, to = "UTF-8", sub = "")
        bad <- str_detect(str_to_lower(lvls), dk_regex) |
          str_detect(lvls, "^\\(88\\)|^\\(99\\)|^\\(999\\)|^\\(9999\\)")
        lvls[bad] <- NA_character_
        levels(.) <- lvls
        droplevels(.)
      })
    ) |>
    mutate(across(where(is.numeric), ~ ifelse(. %in% sentinels, NA_real_, .)))
}

dat <- cmps_latino |>
  mutate(
    voter_fraud_belief = case_when(
      !is.na(Q48) ~ as.character(Q48),
      !is.na(Q49) ~ as.character(Q49),
      !is.na(Q50) ~ as.character(Q50),
      TRUE        ~ NA_character_
    ),
    jan6_fraud_belief = case_when(
      !is.na(Q51) ~ as.character(Q51),
      !is.na(Q52) ~ as.character(Q52),
      TRUE        ~ NA_character_
    )
  ) |>
  mutate(
    voter_fraud_belief = factor(voter_fraud_belief),
    jan6_fraud_belief = factor(jan6_fraud_belief)
  )

dat <- sweep_na(dat)

if ("Q43" %in% names(dat)) {
  dat <- dat |>
    mutate(
      Q43 = if_else(
        str_detect(as.character(Q43), "^\\(6\\)"),
        factor(NA_character_),
        Q43
      ) |> droplevels()
    )
}

miss_rates <- dat |>
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) |>
  pivot_longer(everything(), names_to = "variable", values_to = "pct_missing") |>
  arrange(desc(pct_missing))

exclude_admin <- c(
  ".row_id", "RECORD", "UUID", "DATE", "START_DATE",
  "MAIN_ETH_QUOTA",
  "S2_RACER1", "S2_RACER2", "S2_RACER3", "S2_RACER4",
  "S2_RACER5", "S2_RACER6", "S2_RACER7", "S2_RACER8", "S2_RACE_PRIME",
  "S2_HISPANICR1", "S2_HISPANICR2", "S2_HISPANICR3",
  "S2_HISPANICR4", "S2_HISPANICR5",
  "Q48_50SPLIT", "Q51_52SPLIT", "Q131SPLIT", "Q193SPLIT", "Q199SPLIT",
  "Q257SPLIT", "Q258SPLIT", "Q282SPLIT", "Q285SPLIT", "Q287SPLIT",
  "Q347SPLIT", "Q364SPLIT",
  "Q323BLWSPLIT", "Q323APISPLIT", "Q323NINHSPLIT",
  "LONGSHORTSPLIT", "SPLITQ417_Q424", "SPLITQ517_Q520",
  "SPLITQ634_Q638", "SPLITQ693_Q700", "SPLITQ716", "SPLITQ799", "SPLITA",
  "BLWNAMESPLIT", "BLWETHSPLIT", "BLWGENDERSPLIT", "BLWSTATEMENTSPLIT",
  "APINAMESPLIT", "APIETHSPLIT", "APIGENDERSPLIT", "APISTATEMENTSPLIT",
  "NINHNAMESPLIT", "NINHETHSPLIT", "NINHGENDERSPLIT", "NINHSTATEMENTSPLIT",
  "Q340_RACE_1", "Q340_RACE_2", "Q340_GENDER_1", "Q340_GENDER_2",
  "Q340_EXP_1", "Q340_EXP_2", "Q340_VOUCHER_1", "Q340_VOUCHER_2",
  "Q340_EVICT_1", "Q340_EVICT_2", "Q340_BUILD_1", "Q340_BUILD_2",
  "Q340_GUN_1", "Q340_GUN_2",
  "TABLE1_HID", "TABLE2_HID", "AGE1_HID", "AGE2_HID",
  "PROF1_HID", "PROF2_HID", "POLEXP1_HID", "POLEXP2_HID",
  "RAND_POS1", "RAND_POS2", "RAND_SCORE1", "RAND_SCORE2",
  "Q315_TEXT", "Q607TEXT", "Q3B_PLURAL", "Q294",
  "C2PID", "C3PID", "Q315",
  names(dat)[str_detect(names(dat), "OE$|_TEXT$")],
  "S15", "Q259", "Q426R1", "Q426R2", "Q772", "Q776",
  "Q778R1", "Q781", "Q789R1",
  "VOS", "VBROWSER", "VMOBILEDEVICE", "VMOBILEOS",
  "SCREENTIMESCM6T", "QTIME",
  "WEIGHT", "OS_WEIGHT", "BASELINE",
  "MENA_MUSLIM", "HAWAII_PI", "LGBTQ", "AFROLAT", "BLACKIMM", "NATIVEAM",
  "Q12", "Q13", "S3B",
  ".is_male", ".is_female"
)

exclude_highmiss <- miss_rates |>
  filter(pct_missing > 50) |>
  pull(variable)

exclude_experimental <- c(
  names(dat)[str_detect(names(dat), "^Q195|^Q197|^Q323|^Q340A|^Q340B")],
  "Q196", "Q198", "Q199", "Q200", "Q201", "Q272", "Q340"
)

exclude_tautological <- c(
  "Q14",
  "Q2R1", "Q2R2", "Q2R3", "Q2R4", "Q2R5",
  "Q2R6", "Q2R7", "Q2R8", "Q2R9", "Q2R10",
  "Q15",
  "Q504R11", "Q504R12",
  "Q614",
  "Q96R10", "Q96R11",
  "Q98R8", "Q98R9",
  "Q575_Q581R2",
  "Q575_Q581R7",
  # Direct Trump-approval items (added 2026-04-20)
  "Q130",                  # COVID: Trump not to blame
  "Q54",                   # Jan 6: Trump incited/no connection
  "Q55",                   # Jan 6: Republican Congress undermining/protecting democracy
  "Q56",                   # Jan 6 rioters: White supremacists / racism framing
  "Q684",                  # Belief: ineligible voters cast ballots
  "voter_fraud_belief",    # synthesized from Q48/Q49/Q50
  "jan6_fraud_belief"      # synthesized from Q51/Q52
  # Retained: Q53 (Jan 6 event characterization: protest that went too far vs insurrection)
  # Q53 is the least tautological Jan 6 item and is kept as the single representative
)

exclude_partisan <- c(
  "Q21", "Q22", "Q23", "Q24", "Q25", "Q26",
  "Q43",
  "Q225R10", "Q225R11",
  paste0("Q27R", 1:19),
  paste0("Q28R", 1:19),
  "Q701_Q705R1", "Q701_Q705R2", "Q701_Q705R3",
  "Q701_Q705R4", "Q701_Q705R5",
  "Q743",
  # Top-20 audit addition (2026-04-21, parallel to 2016 L368 / 2024 Q174):
  # Q1R17 asks whether "stopping the left-wing Democratic agenda" is a priority.
  # The item wording explicitly names the Democratic Party, making it a
  # partisan-ID leak rather than a policy-preference attitude. Moved to
  # exclude_partisan so T3 (Ring 1 Non-Partisan) does not include it.
  "Q1R17"
)

exclude_split_originals <- c("Q48", "Q49", "Q50", "Q51", "Q52")

exclude_all <- unique(c(
  exclude_admin,
  exclude_highmiss,
  exclude_experimental,
  exclude_split_originals,
  "trump_vote"
))

feature_pool <- setdiff(names(dat), exclude_all)

message(
  "Preprocessing complete: ", length(feature_pool),
  " candidate features before split-specific imputation and encoding."
)
```

# Train/Test Split and Imputation

```{r}
#| label: split-impute
#| include: false

set.seed(SEED_SPLIT)
train_idx <- caret::createDataPartition(dat$trump_vote, p = 1 - PROP_TEST, list = FALSE)[, 1]

split_obj <- prepare_split_matrices(
  df = dat,
  feature_pool = feature_pool,
  train_idx = train_idx,
  weights_full = weights_full,
  remove_first_dummy = REMOVE_FIRST_DUMMY,
  impute_seed = SEED_IMPUTE
)

x_train_imp <- split_obj$x_train_imp
x_test_imp  <- split_obj$x_test_imp
y_train     <- split_obj$y_train
y_test      <- split_obj$y_test
w_train     <- split_obj$w_train
w_test      <- split_obj$w_test
is_male_train   <- split_obj$is_male_train
is_female_train <- split_obj$is_female_train
is_male_test    <- split_obj$is_male_test
is_female_test  <- split_obj$is_female_test

message(
  "Split complete: ", nrow(x_train_imp), " train / ", nrow(x_test_imp), " test, ",
  ncol(x_train_imp), " features. OHE align: +", split_obj$n_missing_cols,
  " test columns, dropped ", split_obj$n_extra_cols, " extra test columns."
)
```

```{r}
#| label: tbl-split
#| tbl-cap: "Train/Test Split Summary."

tibble(
  Set = c("Training", "Test"),
  N = c(nrow(x_train_imp), nrow(x_test_imp)),
  `Trump %` = round(c(mean(y_train), mean(y_test)) * 100, 1),
  Features = c(ncol(x_train_imp), ncol(x_test_imp))
) |>
  latex_tbl(font_size = 9)
```

# Tier Construction and Partisan Leakage Audit

```{r}
#| label: tiers
#| include: false

tier_obj <- build_tier_sets(x_train_imp, x_test_imp, exclude_tautological, exclude_partisan)

train_t1 <- tier_obj$train_t1
train_t2 <- tier_obj$train_t2
train_t3 <- tier_obj$train_t3
test_t1  <- tier_obj$test_t1
test_t2  <- tier_obj$test_t2
test_t3  <- tier_obj$test_t3
t2_drop  <- tier_obj$t2_drop
t3_drop  <- tier_obj$t3_drop
t2_cols  <- tier_obj$t2_cols
t3_cols  <- tier_obj$t3_cols

# Gender subgroup label vectors are retained in the save bundle so that the
# separate subgroup QMD can reconstruct slices without refitting the main pipeline.
# All subgroup modelling is handled in CMPS_2020_IFA_gender_subgroups.qmd.

message(
  "Tiers: T1=", ncol(train_t1), " T2=", ncol(train_t2), " T3=", ncol(train_t3)
)
```

```{r}
#| label: tbl-tiers
#| tbl-cap: "Progressive Exclusion Tiers."

tibble(
  Tier = c("1: Full", "2: Non-Tautological", "3: Non-Partisan (Ring 1)"),
  `Train N` = c(nrow(train_t1), nrow(train_t2), nrow(train_t3)),
  `Test N` = c(nrow(test_t1), nrow(test_t2), nrow(test_t3)),
  Features = c(ncol(train_t1), ncol(train_t2), ncol(train_t3)),
  Excluded = c(0, length(t2_drop), length(t3_drop))
) |>
  latex_tbl(font_size = 9, scale_down = TRUE)
```

# Model Fitting

```{r}
#| label: models
#| include: false

message("Fitting Tier 1 model...")
rf_t1 <- fit_rf(train_t1, y_train, SEED_RF_T1, survey_w = w_train)
message("Fitting Tier 2 model...")
rf_t2 <- fit_rf(train_t2, y_train, SEED_RF_T2, survey_w = w_train)
message("Fitting Tier 3 model...")
rf_t3 <- fit_rf(train_t3, y_train, SEED_RF_T3, survey_w = w_train)

auc_t1 <- get_auc(rf_t1, test_t1, y_test)
auc_t2 <- get_auc(rf_t2, test_t2, y_test)
auc_t3 <- get_auc(rf_t3, test_t3, y_test)

message(
  "AUC: T1=", round(auc_t1, 3),
  " T2=", round(auc_t2, 3),
  " T3=", round(auc_t3, 3)
)
```

# Benchmark Models

Elastic net logistic regression and XGBoost are fitted on the same Tier 3 train/test split to confirm that findings are a property of the data rather than an artifact of Random Forest.

```{r}
#| label: benchmarks
#| include: false

message("Fitting elastic net benchmark...")
x_train_mat <- as.matrix(train_t3)
x_test_mat  <- as.matrix(test_t3)

n0 <- sum(y_train == 0)
n1 <- sum(y_train == 1)
n_total <- n0 + n1
glm_weights <- ifelse(y_train == 1, n_total / (2 * n1), n_total / (2 * n0))
glm_weights <- glm_weights * w_train

set.seed(SEED_RF_T3)
cv_enet <- glmnet::cv.glmnet(
  x = x_train_mat,
  y = y_train,
  family = "binomial",
  alpha = 0.5,
  weights = glm_weights,
  nfolds = 5,
  type.measure = "auc"
)

pred_enet <- predict(cv_enet, newx = x_test_mat, s = "lambda.min", type = "response")[, 1]
auc_enet  <- as.numeric(pROC::auc(pROC::roc(y_test, pred_enet, quiet = TRUE)))

message("Fitting XGBoost benchmark...")
scale_pos <- n0 / n1

set.seed(SEED_RF_T3)
xgb_train <- xgboost::xgb.DMatrix(data = x_train_mat, label = y_train, weight = w_train)
xgb_test  <- xgboost::xgb.DMatrix(data = x_test_mat, label = y_test)

xgb_model <- xgboost::xgb.train(
  params = list(
    objective = "binary:logistic",
    eval_metric = "auc",
    max_depth = 6,
    eta = 0.1,
    subsample = 0.8,
    colsample_bytree = 0.8,
    scale_pos_weight = scale_pos
  ),
  data = xgb_train,
  nrounds = 300,
  verbose = 0
)

pred_xgb <- predict(xgb_model, xgb_test)
auc_xgb  <- as.numeric(pROC::auc(pROC::roc(y_test, pred_xgb, quiet = TRUE)))

message(
  "Benchmark AUCs: Elastic Net=", round(auc_enet, 3),
  " XGBoost=", round(auc_xgb, 3),
  " Random Forest=", round(auc_t3, 3)
)
```

```{r}
#| label: tbl-benchmarks
#| tbl-cap: "Tier 3 AUC Across Learners."

tibble(
  Learner = c("Random Forest", "Elastic Net Logistic", "XGBoost"),
  `Test AUC` = round(c(auc_t3, auc_enet, auc_xgb), 3)
) |>
  latex_tbl(font_size = 9)
```

# SHAP Computation

```{r}
#| label: shap
#| include: false

message("Computing SHAP values for Tier 1...")
sv_t1 <- compute_shap(rf_t1, train_t1, test_t1)
imp_t1 <- wmean_abs_shap(sv_t1$S, w_test)

message("Computing SHAP values for Tier 2...")
sv_t2 <- compute_shap(rf_t2, train_t2, test_t2)
imp_t2 <- wmean_abs_shap(sv_t2$S, w_test)

message("Computing SHAP values for Tier 3...")
sv_t3 <- compute_shap(rf_t3, train_t3, test_t3)
imp_t3 <- wmean_abs_shap(sv_t3$S, w_test)
```

# Feature Correlation Structure

When features are correlated, SHAP distributes predictive credit across correlated items in ways that are mathematically principled but may not map onto standalone constructs. The correlation matrix below documents the pairwise relationships among the top features.

```{r}
#| label: fig-corr
#| fig-width: 9
#| fig-height: 8
#| fig-cap: "Pairwise Pearson correlations among top 20 Tier 3 features (test set)."

top20_names <- names(sort(imp_t3, decreasing = TRUE))[1:min(20, length(imp_t3))]
top20_data  <- test_t3[, top20_names, drop = FALSE]
top20_labels <- label_feature(top20_names)
names(top20_data) <- stringr::str_trunc(top20_labels, 35)

cor_mat <- cor(top20_data, use = "pairwise.complete.obs")

ggcorrplot::ggcorrplot(
  cor_mat,
  type = "lower",
  lab = TRUE,
  lab_size = 2.5,
  tl.cex = 8,
  colors = c("#B22222", "white", "#1D3557"),
  title = "Correlation Structure: Top 20 Non-Partisan Features"
) +
  theme(plot.title = element_text(size = 12))
```

# Bootstrap Confidence Intervals

The bootstrap resamples the training set `r N_BOOT` times, refitting a Random Forest and computing TreeSHAP on the fixed test set each time. This isolates model instability from evaluation-set composition variance. Each feature's 95% CI is the 2.5th--97.5th percentile of the bootstrap distribution of $E_w[|\phi_j|]$.

```{r}
#| label: boot-t3
#| include: false

message("=== Bootstrap: Tier 3 ===")
boot_t3 <- run_bootstrap(train_t3, test_t3, y_train, y_test, w_train, w_test,
                         SEED_BOOT,           label = "T3")

message("=== Bootstrap: Tier 1 ===")
boot_t1 <- run_bootstrap(train_t1, test_t1, y_train, y_test, w_train, w_test,
                         SEED_BOOT + 1000L,   label = "T1")

message("=== Bootstrap: Tier 2 ===")
boot_t2 <- run_bootstrap(train_t2, test_t2, y_train, y_test, w_train, w_test,
                         SEED_BOOT + 2000L,   label = "T2")
```

# Results Assembly

```{r}
#| label: results
#| include: false

build_results <- function(point_imp, boot_obj, sigma_k = SIGMA_THRESHOLD) {
  out <- tibble(
    feature = names(point_imp),
    label   = label_feature(names(point_imp)),
    E_w_phi = unname(point_imp),
    rank    = rank(-unname(point_imp), ties.method = "average")
  ) |>
    arrange(rank)

  if (!is.null(boot_obj)) {
    bm <- boot_obj$imp
    br <- boot_obj$rank
    out <- out |>
      mutate(
        ci_lo = apply(bm, 2, quantile, probs = 0.025, na.rm = TRUE)[feature],
        ci_hi = apply(bm, 2, quantile, probs = 0.975, na.rm = TRUE)[feature],
        sigma_j = apply(br, 2, sd, na.rm = TRUE)[feature],
        finding = (ci_lo > 0) & (sigma_j <= sigma_k)
      )
  }
  out
}

results_t3 <- build_results(imp_t3, boot_t3)
results_t1 <- build_results(imp_t1, boot_t1)
results_t2 <- build_results(imp_t2, boot_t2)

# Tier 3 test-set probabilities retained for the reporting QMD
pred_t3 <- predict_prob(rf_t3, test_t3)

# Harmonized AUC summary: point AUC plus bootstrap 95% CIs for all three tiers
auc_summary_final <- tibble(
  tier       = c("T1", "T2", "T3"),
  auc_point  = c(auc_t1, auc_t2, auc_t3),
  boot_mean  = c(mean(boot_t1$auc), mean(boot_t2$auc), mean(boot_t3$auc)),
  ci_lo      = c(quantile(boot_t1$auc, 0.025, na.rm = TRUE),
                 quantile(boot_t2$auc, 0.025, na.rm = TRUE),
                 quantile(boot_t3$auc, 0.025, na.rm = TRUE)),
  ci_hi      = c(quantile(boot_t1$auc, 0.975, na.rm = TRUE),
                 quantile(boot_t2$auc, 0.975, na.rm = TRUE),
                 quantile(boot_t3$auc, 0.975, na.rm = TRUE)),
  n_success  = c(boot_t1$n_success, boot_t2$n_success, boot_t3$n_success)
)
```

# Threshold Sensitivity

The stability threshold $k$ is a pre-specified reporting rule, not a universal standard. The table below shows which features qualify as stable findings at $k = 2.0$, $k = 3.0$, and $k = 4.0$, holding the magnitude criterion ($\text{ci\_lo} > 0$) constant throughout.

```{r}
#| label: tbl-sensitivity
#| tbl-cap: "Threshold Sensitivity: Stable Features at k = 2.0, 3.0, and 4.0."

sensitivity <- results_t3 |>
  filter(ci_lo > 0) |>
  slice_head(n = 20) |>
  transmute(
    Rank = as.integer(rank),
    Feature = label,
    sigma_j = round(sigma_j, 2),
    `k=2.0` = ifelse(sigma_j <= 2.0, "\u2713", ""),
    `k=3.0` = ifelse(sigma_j <= 3.0, "\u2713", ""),
    `k=4.0` = ifelse(sigma_j <= 4.0, "\u2713", "")
  )

sensitivity |>
  latex_tbl(font_size = 8, scale_down = TRUE)
```

# Repeated-Resampling Robustness

The bootstrap assesses training-sample sensitivity conditional on one fixed test partition. To confirm the feature architecture is not an artifact of the particular split, five independent 75/25 partitions are drawn and the full Tier 3 RF + SHAP pipeline is run on each. Spearman rank correlations of feature importance across all pairs test whether the same features surface regardless of the split.

```{r}
#| label: resample-robustness
#| include: false

N_RESAMPLE_SPLITS <- if (FAST_DEV) 2L else 5L
resample_seeds <- SEED_SPLIT + seq(100L, by = 100L, length.out = N_RESAMPLE_SPLITS)

resample_imps <- list()

for (s_i in seq_len(N_RESAMPLE_SPLITS)) {
  message("=== Resampling robustness: split ", s_i, " / ", N_RESAMPLE_SPLITS, " ===")

  result <- tryCatch({
    set.seed(resample_seeds[s_i])
    rs_idx <- caret::createDataPartition(dat$trump_vote, p = 1 - PROP_TEST, list = FALSE)[, 1]

    rs_split <- prepare_split_matrices(
      df = dat,
      feature_pool = feature_pool,
      train_idx = rs_idx,
      weights_full = weights_full,
      remove_first_dummy = REMOVE_FIRST_DUMMY,
      impute_seed = SEED_IMPUTE + resample_seeds[s_i]
    )

    rs_tiers <- build_tier_sets(rs_split$x_train_imp, rs_split$x_test_imp,
                                exclude_tautological, exclude_partisan)

    rs_rf <- fit_rf(rs_tiers$train_t3, rs_split$y_train,
                    seed = resample_seeds[s_i], survey_w = rs_split$w_train)
    rs_sv <- compute_shap(rs_rf, rs_tiers$train_t3, rs_tiers$test_t3)
    wmean_abs_shap(rs_sv$S, rs_split$w_test)
  }, error = function(e) {
    message("  Split ", s_i, " failed entirely: ", conditionMessage(e), " — skipping")
    NULL
  })

  resample_imps[[s_i]] <- result
}

resample_imps <- Filter(Negate(is.null), resample_imps)
if (length(resample_imps) < 2) {
  warning("Fewer than 2 resample splits succeeded — skipping robustness analysis.")
  min_rho  <- NA_real_
  mean_rho <- NA_real_
  rho_mat  <- matrix(NA_real_, 1, 1, dimnames = list("Split_1", "Split_1"))
} else {
  common_features <- Reduce(intersect, lapply(resample_imps, names))
  imp_mat <- sapply(resample_imps, function(imp) imp[common_features])

  n_splits <- ncol(imp_mat)
  rho_mat <- matrix(NA_real_, n_splits, n_splits,
                    dimnames = list(paste0("Split_", 1:n_splits), paste0("Split_", 1:n_splits)))
  for (i in 1:(n_splits - 1)) {
    for (j in (i + 1):n_splits) {
      rho_mat[i, j] <- cor(imp_mat[, i], imp_mat[, j], method = "spearman")
      rho_mat[j, i] <- rho_mat[i, j]
    }
  }
  diag(rho_mat) <- 1.0
  min_rho  <- min(rho_mat[upper.tri(rho_mat)])
  mean_rho <- mean(rho_mat[upper.tri(rho_mat)])

  message("Resampling robustness: min rho = ", round(min_rho, 3),
          " mean rho = ", round(mean_rho, 3))
}
```

```{r}
#| label: tbl-resample
#| tbl-cap: !expr paste0("Spearman Rank Correlations Across ", N_RESAMPLE_SPLITS, " Independent Splits (min = ", round(min_rho, 3), ", mean = ", round(mean_rho, 3), ").")

as.data.frame(round(rho_mat, 3)) |>
  tibble::rownames_to_column("Split") |>
  latex_tbl(font_size = 8, scale_down = TRUE)
```

# Main Results

## Model Performance

```{r}
#| label: tbl-auc
#| tbl-cap: "AUC Across the Progressive Exclusion Framework."

tibble(
  Model = c("Tier 1 (Full)", "Tier 2 (Non-Tautological)",
            "Tier 3 (Non-Partisan)"),
  `Test AUC` = round(c(auc_t1, auc_t2, auc_t3), 3),
  `Drop from T1` = c(NA, round(auc_t1 - auc_t2, 3), round(auc_t1 - auc_t3, 3)),
  `Boot Mean AUC` = c(NA, NA, round(mean(boot_t3$auc), 3)),
  `95% CI` = c("---", "---", auc_ci(boot_t3$auc))
) |>
  latex_tbl(font_size = 9, scale_down = TRUE)
```

## Stable Non-Partisan Features

Features are reported as stable findings under a conjunctive decision rule pre-specified before examining results: the bootstrap 95% CI lower bound for $E_w[|\phi_j|]$ must exceed zero (magnitude criterion) AND $\sigma_j \leq$ `r SIGMA_THRESHOLD` (rank stability criterion). The first condition ensures the feature's importance is bounded away from nothing across `r N_BOOT` training resamples. The second ensures its ordinal position in the importance ranking is stable enough to support comparative claims.

```{r}
#| label: tbl-stable
#| tbl-cap: !expr paste0("Stable Ring 1 Non-Partisan Features (ci_lo > 0, sigma_j <= ", SIGMA_THRESHOLD, ").")

results_t3 |>
  filter(finding) |>
  transmute(
    Rank = as.integer(rank),
    Feature = label,
    Code = feature,
    `E_w[|phi|]` = round(E_w_phi, 5),
    `CI lo` = round(ci_lo, 5),
    `CI hi` = round(ci_hi, 5),
    sigma_j = round(sigma_j, 2)
  ) |>
  latex_tbl(font_size = 8, scale_down = TRUE)
```

## Figure 1. Feature Importance (Beeswarm)

```{r}
#| label: fig-beeswarm
#| fig-width: 9
#| fig-height: 8

sv_t3_labeled <- sv_t3
colnames(sv_t3_labeled$S) <- label_feature(colnames(sv_t3_labeled$S))
colnames(sv_t3_labeled$X) <- label_feature(colnames(sv_t3_labeled$X))

shapviz::sv_importance(sv_t3_labeled, kind = "beeswarm", max_display = 15L, alpha = 0.45) +
  labs(
    title = "Ring 1 Non-Partisan Predictors of Latino Trump Support (2020)",
    subtitle = paste0("Test set | N = ", nrow(test_t3), " | partisan alignment removed")
  ) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank())
```

## Figure 2. Bootstrap Confidence Intervals

```{r}
#| label: fig-ci
#| fig-width: 9
#| fig-height: 7

ci_data <- results_t3 |>
  slice_head(n = 20) |>
  mutate(
    Feature = stringr::str_trunc(label, 48),
    Feature = forcats::fct_reorder(Feature, -rank)
  )

ggplot(ci_data, aes(x = E_w_phi, y = Feature, color = finding)) +
  geom_point(size = 2.5) +
  geom_errorbarh(aes(xmin = ci_lo, xmax = ci_hi), height = 0.3) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  scale_color_manual(
    values = c("TRUE" = "#1D3557", "FALSE" = "grey65"),
    labels = c(
      "TRUE" = paste0("Stable (ci_lo > 0, ", "\u03c3", "\u2c7c", " \u2264 ", SIGMA_THRESHOLD, ")"),
      "FALSE" = "Does not meet both criteria"
    ),
    name = NULL
  ) +
  labs(
    title = "Bootstrap CIs for Ring 1 Non-Partisan Features",
    subtitle = paste0("Top 20 | B = ", boot_t3$n_success,
                      " | Decision rule: ci_lo > 0 and ", "\u03c3", "\u2c7c", " \u2264 ", SIGMA_THRESHOLD),
    x = "Weighted mean |SHAP|",
    y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

## Figure 3. SHAP Dependence (Top Feature)

```{r}
#| label: fig-dependence
#| fig-width: 8
#| fig-height: 5

top_feature <- results_t3$feature[1]
shapviz::sv_dependence(sv_t3, v = top_feature, alpha = 0.4) +
  labs(
    title = paste0("SHAP Dependence: ", label_feature(top_feature)),
    y = "SHAP value"
  ) +
  theme_minimal(base_size = 11) +
  theme(panel.grid.minor = element_blank())
```

# Save Results Bundle

All model objects, SHAP matrices, bootstrap results, tier metadata, and summary tables are saved to a single `.rds` file. The reporting QMD loads this file to generate all manuscript figures and tables without re-running the model.

```{r}
#| label: save-rds
#| include: false

ifa_results <- list(
  render_date = Sys.Date(),
  r_version = R.version.string,
  params = list(
    n_boot = N_BOOT,
    n_trees = N_TREES,
    prop_test = PROP_TEST,
    fast_dev = FAST_DEV,
    sigma_threshold = SIGMA_THRESHOLD,
    use_survey_weights = USE_SURVEY_WEIGHTS,
    remove_first_dummy = REMOVE_FIRST_DUMMY
  ),
  seeds = list(
    split = SEED_SPLIT,
    impute = SEED_IMPUTE,
    rf_t1 = SEED_RF_T1,
    rf_t2 = SEED_RF_T2,
    rf_t3 = SEED_RF_T3,
    rf_men = SEED_RF_MEN,
    rf_women = SEED_RF_WOMEN,
    boot = SEED_BOOT,
    boot_t1 = SEED_BOOT + 1000L,
    boot_t2 = SEED_BOOT + 2000L
  ),
  # Subgroup label vectors retained for downstream gender-subgroup QMD.
  is_male_train   = is_male_train,
  is_male_test    = is_male_test,
  is_female_train = is_female_train,
  is_female_test  = is_female_test,
  raw_data_file = basename(RAW_DATA_FILE),
  raw_data_md5 = unname(tools::md5sum(RAW_DATA_FILE)),

  n_latino = nrow(cmps_latino),
  n_trump = sum(cmps_latino$trump_vote),
  train_idx = train_idx,
  n_train = nrow(x_train_imp),
  n_test = nrow(x_test_imp),

  feature_labels = feature_labels,
  all_features = names(x_train_imp),
  feature_pool = feature_pool,
  t2_drop = t2_drop,
  t3_drop = t3_drop,
  t2_cols = t2_cols,
  t3_cols = t3_cols,
  n_features_t1 = ncol(train_t1),
  n_features_t2 = ncol(train_t2),
  n_features_t3 = ncol(train_t3),

  auc = c(t1 = auc_t1, t2 = auc_t2, t3 = auc_t3),
  auc_enet = auc_enet,
  auc_xgb = auc_xgb,

  # Tier 3 feature matrices retained so the gender-subgroup QMD can
  # reconstruct men/women slices and refit Models 4 and 5 without rerunning
  # the full pipeline.
  train_t3 = train_t3,
  test_t3  = test_t3,
  y_train  = y_train,
  y_test   = y_test,
  w_train  = w_train,

  sv_t1 = sv_t1,
  sv_t2 = sv_t2,
  sv_t3 = sv_t3,

  imp_t1 = imp_t1,
  imp_t2 = imp_t2,
  imp_t3 = imp_t3,

  boot_t1 = boot_t1,
  boot_t2 = boot_t2,
  boot_t3 = boot_t3,

  exclude_tautological = exclude_tautological,
  exclude_partisan = exclude_partisan,

  sigma_threshold = SIGMA_THRESHOLD,
  cor_top20 = cor_mat,
  sensitivity = sensitivity,
  resample_rho_mat = rho_mat,
  resample_min_rho = min_rho,
  resample_mean_rho = mean_rho,

  results_t1 = results_t1,
  results_t2 = results_t2,
  results_t3 = results_t3,

  pred_t3 = pred_t3,
  auc_summary_final = auc_summary_final,

  w_test = w_test
)

rds_path <- file.path(PROJECT_ROOT, "ifa_results.rds")
saveRDS(ifa_results, rds_path)
message("Results saved to: ", rds_path, " (", round(file.size(rds_path) / 1e6, 1), " MB)")
```

The results bundle has been saved to `ifa_results.rds` (`r round(file.size(file.path(PROJECT_ROOT, "ifa_results.rds")) / 1e6, 1)` MB). Load this file in the reporting QMD with `readRDS("ifa_results.rds")`.

# Session Information

```{r}
#| label: session-info

sessioninfo::session_info()
```
